Inference of seasonal forcings on reproduction numbers during epidemics

### 2020-04-18 PaperDraft.Rmd

For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing

$$
p_{ij} = \frac{w(t_i - t_j)}{\sum_{i \neq k} }
$$

@@

[TK: Beta values]
Fitting our Bayesian model, which extends the method of Thompson et. al, values
of beta were estimated for temperature, relative humidity, Gross Domestic
Product (GDP) and population density, respectively. The mean estimate of $\beta^T$ is

@@

in the overall model for $R$

$$ R_i(t) = \exp\left( \theta(t) + \lambda_i + \mathbf{\beta}^T x_i(t) + \varepsilon_i(t) \right)$$
Note that each feature's data is standardized to mean 0 and variance 1.

@@

### Climate-Independent Basic R, by Date

@@

### Estimated R in Cities by Date

\begin{center}
\includegraphics[width=.49\textwidth]{plots/plot6.png}
\includegraphics[width=.49\textwidth]{plots/plot7.png}
\includegraphics[width=.49\textwidth]{plots/plot8.png}
\includegraphics[width=.49\textwidth]{plots/plot9.png}
\includegraphics[width=.49\textwidth]{plots/plot10.png}
\end{center}

@@

@@

- You want to minimize the value episilon.

>>>>

For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$. This method is outlined in detail in Wallinga and Teunis (2004)[8].

@@

Fitting our Bayesian model, which extends the method of Thompson et. al, values
of beta were estimated for temperature, relative humidity, Gross Domestic
Product (GDP) and population density, respectively. The mean estimate of $\beta^T$ is

@@

in the overall model for $R$:

Note that each feature's data is standardized to mean 0 and variance 1.

@@

### Estimates of $\theta$, transformed to climate-independent $R$ by date

@@

### Estimated R in Various Provinces

\begin{center}
\includegraphics[width=.49\textwidth]{plots/plot6.png}
\includegraphics[width=.49\textwidth]{plots/plot7.png}
\includegraphics[width=.49\textwidth]{plots/plot8.png}
\includegraphics[width=.49\textwidth]{plots/plot9.png}
\end{center}

@@

In many cities, our estimated $R_0$ values are extremely high. This is because
those cities reported a dramatic increase in cases, and our estimated $R_0$
values are indeed consistent with the data. However, we will discuss in the
interpretation section the limits of our approach, and why we believe this may
not match the ground truth.

@@

This histogram shows all samples of all entries in the $\varepsilon$ matrix.
Generally, we prefer a lower average spread of $\varepsilon$, because it means
that more of the variation in reproduction number can be predicted from the
seasonal factors, location, and time alone. Unfortunately, we found the observed
spread of $\varepsilon$ is quite high, indicating that there is substantial
variance in reproductive numbers inferred from the data which cannot be
explained by modeled features.
